Human Capital Model

### 2016-02-01 Tex/paper.tex
....

\subsection{Restatement of the Problem}
\label{sec:restatement-of-the-problem}
We are required to build a mathematical model to solve the problem of heating water in bathtub. Thus we have some subproblems:
\begin{itemize}
\item Build a basic model that demonstrates the change of the temperature in the bathtub in space and time without other interventions.
\item Figure out the influence of parameters such as the shape of the bathtub or the motions made by the person or so.
\item Propose a strategy to keep the temperature as even as possible.
\end{itemize}
In the first step, we build the simplest model. The shape of the bathtub is a cuboid, and the person will stay still. In the second step, the person moves slowly and the shape of the bathtub changes. In the third step, we develop the conclusion of the strategy.

@@

Definitions of symbols employed in this paper are listed in Table \ref{1_t} and Table \ref{2_t}.
\begin{table}
\begin{tabular}{l|l}
  Constant & Description \\
  \hline
  $k$            &Thermal conductivity of water \\
  $\rho$         &Density of water \\
  ${\rho}_a$     &Density of air \\
  ${\gamma}_1$   &Heat exchange coefficient between water and air \\
  ${\gamma}_2$   &Heat exchange coefficient between water and person \\
  ${\gamma}_3$   &Heat exchange coefficient between water and bathtub \\
  $C_w$          &Specific heat capacity of water \\
  $C_a$          &Specific heat capacity of air \\
  $V$            &Total volume of water \\
  $V_a$          &Total volume of air above the water \\
  $v$            &Flow velocity of air\\
  $x_s$          &Humidity ratio in saturated air \\
  $x$            &Humidity ratio in the air \\
  $x_0$          &Initial humidity ratio in the air \\
  $A$            &Square of water surface \\
  $V_a$          &Volume of surrounding air \\
  $C_v$          &Evaporation heat of water \\
\end{tabular}
\caption{Constant}\label{1_t}
\end{table}

\begin{table}
\begin{tabular}{l|l}
  Variable & Description \\
  \hline
  $t$            &time \\
  $u$            &Temperature of water \\
  $u$            &Temperature of water when t=0 \\
  $u_1$          &Temperature of air \\
  $u_2$          &Temperature of person's surface \\
  $u_3$          &Temperature of bathtub \\
  $u_h$          &Temperature of hot water from faucet \\
  $v_h$          &Velocity of hot water from faucet \\
  $g$            &Evaporation heat of water\\
\end{tabular}
\caption{Variable}\label{2_t}
\end{table}

@@

In our model, five factors account for heat transfer of water: heat exchange between water and air, heat
exchange between water and the person, heat exchange between water and the bathtub, heat energy of hot water from the
faucet and heat energy of water flow outside. According to thermology, we deduce the equation
of water temperature in the bathtub. The initial-boundary value conditions are deduced according to
Fourier theorem, vaporization equation\cite{1}.

@@

Figure \ref{1_p} shows thermal analysis of water in bathtub.
\begin{figure}[htb]
  \centering
  \includegraphics[width=8cm]{2.pdf}\\
  \caption{Thermal Analysis of Water}\label{1_p}
\end{figure}

First, we assume that the heat change of all the water in a tiny time $\Delta t$ from $t_1$ to $t_2$ is $\Delta Q$.
As shown in Figure 1, heat of water in the bathtub exchanges with air, the person and the bathtub.
In addition, heat changes when the hot water flows in and the excess water flows out.
Thus $\Delta Q$ is divided into five parts, i.e.
\begin{equation}
 \Delta Q=\Delta Q_1+\Delta Q_2-\Delta Q_3-\Delta Q_4
\end{equation}
in which $\Delta Q_1$ is the heat transferred from air, person and bathtub,
$\Delta Q_2$ is the heat of the hot water from the faucet in $\Delta t$,
$\Delta Q_3$ is the heat of the water flows away in $\Delta t$,
$\Delta Q_4$ is the heat loss because of vaporization.

@@

Considering the right side of equation 1,
 Fourier theorem indicates that,
\begin{equation}
 dQ=-k\frac{\partial u}{\partial n}dSdt
\end{equation}
holds for the heat flows through surface with square of dS, as it shows in Figure \ref{2_p}.
\begin{figure}[htb]
  \centering
  \includegraphics[width=8cm]{1.pdf}\\
  \caption{Thermal Analysis}\label{2_p}
\end{figure}
In the equation, k is thermal conductivity, $\frac{\partial u}{\partial n}$ is the derivative of u along the normal line. Thus, after integration of dQ by time and space
\begin{equation}
 \Delta Q_1=\int_{t_1}^{t_2}[{\int_{\Gamma}}k\frac{\partial u}{\partial n}dS]dt
\end{equation}
in which $\Gamma$ is the boundary of $\Omega$,
the heat flows into the water by thermal conduction could be calculated.

@@

Now let's discuss g, evaporation rate per square.
As Figure \ref{3_p} shows, water evaporates into air and air humidity increases, until it is saturate. According to empirical formula\cite{1}, we know that
\begin{equation}
 g=\frac{(25+19v)({x_s}-x)}{3600}
\end{equation}
in which v is the velocity of air, ${x_s}$ is humidity ratio in saturated air, $x$ is unsaturated humidity ratio in air.
Considering the humidity ratio in the air, it indicates
\begin{equation}
 x=x_0+\int \frac{Ag}{\rho V_a}dt
\end{equation}
in which $x_0$ is the initial humidity ratio in the air, $\rho$ is the density of water, $V_a$ is the total volume of surrounding air.
\begin{figure}[htb]
  \centering
  \includegraphics[width=8cm]{3.pdf}\\
  \caption{Evaporation}\label{3_p}
\end{figure}

@@

Boundary value conditions are quite essential in partial differential equations. In our model, water is not a regular object. Thus boundary value condition is complex.

@@

Since bather is temperature-constant, the boundary value condition of ${\Gamma}_1$ and ${\Gamma}_3$ could be considered
as the Dirichlet value condition. That is
\begin{equation}
 u=u_i, (x,y,z)\in {\Gamma}_i, i=2.4
\end{equation}
In the equation, $u_2$ is the surface temperature of the person. We assume it is constant. $u_4$ is the temperature of hot water.

The boundary value condition of ${\Gamma}_1$ and ${\Gamma}_3$ could be considered as the third boundary value condition approximately. That is
\begin{equation}
 k\frac{\partial u}{\partial n}+{{\gamma}_i}u={{\gamma}_i}{u_i}, (x,y,z)\in {\Gamma}_i
\end{equation}
In the equation, i=1 or 3 ,$u_i$ is the temperature of contact material, and 
${\gamma}_i$ is the heat exchange coefficient of ${\Gamma}_i$.

@@

$u_1$, the temperature of air, is time-dependant, because water vapor transfer heat to air. We can calculate $u_1$ by calculating the energy of water vapor evaporating in the air. Referencing the deduction of evaporation in the last section,
\begin{equation}
  u_1=u_0+\frac{C_w\rho}{C_a{{\rho}_a}} (x_s-x_0)(1-e^{-\lambda t})
\end{equation}

@@

\subsection{PDE Solution}
\label{sec:PDE solution}
Since our equations are complex and the region is quite irregular, it's impossible to get a analytical solution. Finite element method is normally adopted to deal with this kind of problem.

We use Partial Differential Equation Toolbox in matlab to solve our equations. Partial Differential Equation Toolbox makes use of finite element method to separate irregular geometry, calculate a relatively precise solution, and get solution of temperature corresponding to different $u_h$ and $v_h$. By comparing temperature distributions, we choose proper parameters as the best strategy.

First, we build a 3D geometry model to simulate a proper-size bathtub, as it shown in Figure \ref{4_p}. Matlab divides the geometry into several elements.
\begin{figure}[htb]
  \centering
  \includegraphics[width=8cm]{4.png}\\
  \caption{3D geometry of water}\label{4_p}
\end{figure}

@@

Our strategy focus on two variables, $v_h$ and $u_h$. We test about fifty pairs of data to find the best strategy. And we list nine representative pairs of data in Table \ref{3_t}.

\begin{table}
\begin{center}
\begin{tabular}{c|c}
\hline
    $v_h kg/s$  &${u_h}\,^{\circ}\mathrm{C}$        \\ \hline
     0.1             & 50              \\ \hline
    0.2             & 50              \\ \hline
     0.3             & 50              \\ \hline
     0.1             & 40              \\ \hline
     0.2             & 40              \\ \hline
     0.3             & 40              \\ \hline
     0.1             & 30              \\ \hline
     0.2             & 30              \\ \hline
     0.3             & 30              \\ \hline
\end{tabular}
\end{center}
\caption{Represenative Data Pairs}\label{3_t}
\end{table}

In each operation, $u_2$ is taken as $33.5\,^{\circ}\mathrm{C}$
and the initial value condition is $35\,^{\circ}\mathrm{C}$.
 We take 120s as time period. The following graphs show the results we get.

Figure \ref{5_p}, Figure \ref{6_p}, Figure \ref{7_p} show the temperature of water at t=240, with $u_h=40$ and $v_h=0.1, 0.2 and 0.3$
\begin{figure}[htb]
  \centering
  \includegraphics[width=8cm]{6-2.pdf}\\
  \caption{$u_h=40$, $v_h=0.1$, t=240}\label{5_p}
\end{figure}

\begin{figure}[htb]
  \centering
  \includegraphics[width=8cm]{10-2.pdf}\\
  \caption{$u_h=40$, $v_h=0.2$, t=240}\label{6_p}
\end{figure}

\begin{figure}[htb]
  \centering
  \includegraphics[width=8cm]{11-2.pdf}\\
  \caption{$u_h=40$,$v_h=0.3$,t=240}\label{7_p}
\end{figure}

>>>>

\subsection{Restatement of the Problem}
\label{sec:restatement-of-the-problem}
We are required to build a mathematical model to solve the problem of heating water in bathtub. Thus we have some subproblems:
\begin{itemize}
\item Build a basic model that demonstrate the change of the temperature in the bathtub in space and time without other intervention.
\item Figure out the influence of parameters such as the shape of the bathtub or the motions made by the person or so.
\item Propose a strategy to keep the temperature as even as possible.
\end{itemize}
In the first step, we build the simplest model. The shape of the bathtub is a cuboid, and the person will stay still. In the second step, the person move slowly and the shape of the bathtub changes. In the third step, we develop the conclusion of the strategy.

@@

Definitions of symbols employed in this paper are listed in
\textbf{Table 1}.
\begin{table}
\begin{tabular}{l|l}
  Constant & Description \\
  \hline
  $k$            &Thermal conductivity of water \\
  $\rho$         &Density of water \\
  ${\rho}_a$     &Density of air \\
  ${\gamma}_1$   &Heat exchange coefficient between water and air \\
  ${\gamma}_2$   &Heat exchange coefficient between water and person \\
  ${\gamma}_3$   &Heat exchange coefficient between water and bathtub \\
  $C_w$          &Specific heat capacity of water \\
  $C_a$          &Specific heat capacity of air \\
  $V$            &Total volume of water \\
  $V_a$          &Total volume of air above the water \\
  $v$            &Flow velocity of air\\
  $x_s$          &Humidity ratio in saturated air \\
  $x$            &Humidity ratio in the air \\
  $x_0$          &Initial humidity ratio in the air \\
  $A$            &Square of water surface \\
  $V_a$          &Volume of surrounding air \\
  $C_v$          &Evaporation heat of water \\
\end{tabular}
\end{table}

\textbf{Table 2}.
\begin{table}
\begin{tabular}{l|l}
  Variable & Description \\
  \hline
  $t$            &time \\
  $u$            &Temperature of water \\
  $u$            &Temperature of water when t=0 \\
  $u_1$          &Temperature of air \\
  $u_2$          &Temperature of person's surface \\
  $u_3$          &Temperature of bathtub \\
  $u_h$          &Temperature of hot water from faucet \\
  $v_h$          &Velocity of hot water from faucet \\
  $g$            &Evaporation heat of water\\
\end{tabular}
\end{table}

@@

In our model, five factors account for heat transfer of water: heat exchange between water and air, heat
exchange between water and the person, heat exchange between water and the bathtub, heat energy of hot water from the
faucet and heat energy of water flow outside. According to thermology, we deduce the equation
of water temperature in the bathtub. The initial-boundary value conditions are deduced according to
Fourier theorem, vaporization equation[ ].

@@

Figure 1 shows thermal analysis of water in bathtub.
In the photo, $W_1$ is heat exchange between the person and water, while $W_2$ is heat from water vaporization. $W_3$ stands for heat exchange between the bathtub and water, and $W_4$ is heat conductivity between water and air. $W_5$ represents heat exchange between the bathtub and air, and $W_6$ is heat increasement with hot in-water.
\begin{figure}[htb]
  \centering
  \includegraphics[width=8cm]{2.pdf}\\
  \caption{Thermal Analysis of Water}\label{1_p}
\end{figure}

First, we assume that the heat change of all the water in a tiny time $\Delta t$ from $t_1$ to $t_2$ is $\Delta Q$.
As shown in Figure 1, heat of water in the bathtub exchanges with air, the person and the bathtub.
In addition, heat changes when the hot water flow in and the excess water flow out.
Thus $\Delta Q$ is divided into five parts, i.e.
\begin{equation}
 \Delta Q=\Delta Q_1+\Delta Q_2-\Delta Q_3-\Delta Q_4
\end{equation}
in which $\Delta Q_1$ is the heat transferred from air, person and bathtub,
$\Delta Q_2$ is the heat of the hot water from the faucet in $\Delta t$,
$\Delta Q_3$ is the heat of the water flow away in $\Delta t$,
$\Delta Q_4$ is the heat loss because of vaporization.

@@

Considering the right side of equation 1.
Fourier theorem indicates that,
\begin{equation}
 dQ=-k\frac{\partial u}{\partial n}dSdt
\end{equation}
holds for the heat flow through surface with square of dS, as it shows in Figure 2.
\begin{figure}[htb]
  \centering
  \includegraphics[width=8cm]{1.pdf}\\
  \caption{Thermal Analysis}\label{2_p}
\end{figure}
In the equation, k is thermal conductivity, $\frac{\partial u}{\partial n}$ is the derivative of u along the normal line. Thus, after integration of dQ by time and space
\begin{equation}
 \Delta Q_1=\int_{t_1}^{t_2}[{\int_{\Gamma}}k\frac{\partial u}{\partial n}dS]dt
\end{equation}
in which $\Gamma$ is the boundary of $\Omega$,
the heat flow into the water by thermal conduction could be calculated.

@@

Now let's discuss g, evaporation rate per square.
As Figure 3 shows, water evaporates into air and air humidity increases, until it is saturate. According to empirical formula in [], we know that
\begin{equation}
 g=\frac{(25+19v)({x_s}-x)}{3600}
\end{equation}
in which v is the velocity of air, ${x_s}$ is humidity ratio in saturated air, $x$ is unsaturated humidity ratio in air.
Considering the humidity ratio in the air, it indicates
\begin{equation}
 x=x_0+\int \frac{Ag}{\rho V_a}dt
\end{equation}
in which $x_0$ is the initial humidity ratio in the air, $\rho$ is the density of water, $V_a$ is the total volume of surrounding air.
\begin{figure}[htb]
  \centering
  \includegraphics[width=8cm]{3.pdf}\\
  \caption{Evaporation}\label{3_p}
\end{figure}

@@

Boundary value conditions are quite essential in partial differential equations. In our model, water is not a regular research object.
Thus boundary value condition is complex.

@@

Since person is temperature-constant, the boundary value condition of ${\Gamma}_1$ and ${\Gamma}_3$ could be considered
as the Dirichlet value condition. That is
\begin{equation}
 u=u_i, (x,y,z)\in {\Gamma}_i, i=2.4
\end{equation}
In the equation, $u_2$ is the surface temperature of the person. We assume it is constant. $u_4$ is the temperature of hot water.

The boundary value condition of ${\Gamma}_1$ and ${\Gamma}_3$ could be considered as the third boundary value condition approximately. That is
\begin{equation}
 k\frac{\partial u}{\partial n}+{{\gamma}_i}u={{\gamma}_i}{u_i}, (x,y,z)\in {\Gamma}_i
\end{equation}
In the equation, i=1 or 3 and $u_i$ is the temperature of contact material,
${\gamma}_i$ is the heat exchange coefficient of ${\Gamma}_i$.

@@

While $u_1$ i, the temperature of air, is time-dependant, because water vapor transfer heat to air. We can calculate $u_1$ by calculating the energy of water vapor evaporated in the air. Referencing the deduction of evaporation in the last section,
\begin{equation}
  u_1=u_0+\frac{C_w\rho}{C_a{{\rho}_a}} (x_s-x_0)(1-e^{-\lambda t})
\end{equation}

@@

\subsection{PDE Solution}
\label{sec:PDE solution}
Since our equations are complex and the region is quite irregular, it's impossible to get a regular solution. Finite element method is normally adopted to deal with this kind of problem.

We use Partial Differential Equation Toolbox in matlab to solve our equations. Partial Differential Equation Toolbox makes use of finite element method to separate irregular geometry and calculate a relatively precise solution, get solution of temperature corresponding to different $u_h$ and $v_h$. By comparing temperature distribution, we choose proper parameters as the best strategy.

First, we build a 3D geometry model to simulate a proper-size bathtub, as it shown in Figure 4. Matlab divided the geometry into several elements.
\begin{figure}[htb]
  \centering
  \includegraphics[width=8cm]{4.png}\\
  \caption{3D geometry of water}\label{4_p}
\end{figure}

@@

Our strategy focus on two variables, $v_h$ and $u_h$. Nine pairs of data are tested to find the best strategy.
\begin{center}
\begin{tabular}{c|c}
\hline
    $v_h kg/s$  &${u_h}\,^{\circ}\mathrm{C}$        \\ \hline
     0.1             & 50              \\ \hline
    0.2             & 50              \\ \hline
     0.3             & 50              \\ \hline
     0.1             & 40              \\ \hline
     0.2             & 40              \\ \hline
     0.3             & 40              \\ \hline
     0.1             & 30              \\ \hline
     0.2             & 30              \\ \hline
     0.3             & 30              \\ \hline
\end{tabular}
\end{center}
In each operation, $u_2$ is taken as $33.5\,^{\circ}\mathrm{C}$
and the initial value condition is $35\,^{\circ}\mathrm{C}$.
 We take 120s as time period. Then the result performs like the following graphs show.

Figure 5, Figure 6, Figure 7 shows the temperature of water at t=240, with $u_h=40$ and $v_h=0.1, 0.2 and 0.3$
\begin{figure}[htb]
  \centering
  \includegraphics[width=8cm]{6-2.pdf}\\
  \caption{$u_h=40$,$v_h=0.1$,t=240}\label{6-2_p}
\end{figure}

\begin{figure}[htb]
  \centering
  \includegraphics[width=8cm]{11-2.pdf}\\
  \caption{$u_h=40$,$v_h=0.3$,t=240}\label{11-2_p}
\end{figure}
